---
title: "A Tour of the Tidyverse"
subtitle: "Data Wrangling and Visualization in R"
author: "Jake Riley, Alice Walsh"
date: "March 28, 2020"
output: 
  powerpoint_presentation:
    reference_doc: reference.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Welcome
*Can use a different ppt template! This is draft!*
Link to RStudioCloud on every slide

## Agenda

## Intro Jake

Link to RStudioCloud on every slide

## Intro Alice

Link to RStudioCloud on every slide

## What is Tidyverse

The tidyverse is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures.

Install the complete tidyverse with:

```{r eval=FALSE, echo=TRUE}
install.packages("tidyverse")
```


[https://www.tidyverse.org](https://www.tidyverse.org)

## Intro

+ what is `%>%` ?

shortcut = ctrl+shift+m

## Schedule

```{r}
schedule <- data.frame(time = c("2:00-2:15",
                                "2:15-2:30",
                                "2:30-2:45",
                                "2:45-3:30",
                                "3:30-3:45",
                                "3:45-4:45",
                                "4:45-5:00"),
                       event = c("Welcome and Intros",
                                 "Intro to data cleaning",
                                 "Break",
                                 "Data cleaning exercises",
                                 "Break",
                                 "Data visualization",
                                 "Goodbye!"))

schedule
```

## Get to RStudioCloud

LINK

