---
title: "A Tour of the Tidyverse"
subtitle: "Data Wrangling and Visualization in R"
author: "Jake Riley, Alice Walsh"
date: "March 28, 2020"
output: 
  powerpoint_presentation:
    reference_doc: reference.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Welcome
*Can use a different ppt template! This is a draft!*


Link to RStudioCloud on every slide

## Agenda
```{r}
schedule <- data.frame(time = c("2:00-2:15",
                                "2:15-2:30",
                                "2:30-2:45",
                                "2:45-3:30",
                                "3:30-3:45",
                                "3:45-4:45",
                                "4:45-5:00"),
                       event = c("Welcome and Intros",
                                 "Intro to data cleaning",
                                 "Break",
                                 "Data cleaning exercises",
                                 "Break",
                                 "Data visualization",
                                 "Goodbye!"))

schedule
```
## Jake



## Alice

Hello! I am a computational biologist that uses R to better understand disease at the molecular level. I work at Bristol Myers Squibb in Princeton, NJ.

github.com/awalsh17

@sciencealice

## What is the Tidyverse?

The tidyverse is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures.

Install the complete tidyverse with:

```{r eval=FALSE, echo=TRUE}
install.packages("tidyverse")
```


[https://www.tidyverse.org](https://www.tidyverse.org)

## Some important intro notes!

What is `%>%` ?

This is the "pipe operator". It is a way to chain multiple operations together.

```{r eval=FALSE, echo=TRUE}
# Without pipe
round(exp(diff(log(x))), 1)
# With pipe
x %>% log() %>%
    diff() %>%
    exp() %>%
    round(1)
```


**keyboard shortcut = ctrl+shift+m**


## Link to RStudioCloud

LINK

