---
title: "Untitled"
author: "Alice Walsh"
date: "3/2/2020"
output: 
  powerpoint_presentation:
    reference_doc: reference.pptx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Backups/Reference slides

## Getting the @#$@! Data into R

1. Best case scenario - Rds file

```{r eval=FALSE, echo=TRUE}
mydata <- readRDS('mydata.Rds')
```

## Getting the @#$@! Data into R

2. Still good scenario - flat text file (csv, tab-delimited)

```{r eval=FALSE, echo=TRUE}
mydata <- read.csv('mydata.csv')
```



Variations on base R read.csv():

+ `readr::read_csv()` 
+ `data.table::fread()`

## Getting the @#$@! Data into R

### What's the difference?   

+ stringsAsFactors
+ "bad" column names
+ others?

## Getting the @#$@! Data into R

### Other data sources

+ database (local or remote)
+ JSON
+ xml

### At the end of the day, we usually want a data frame!!

## First challenge - load our data

Dataset source, [opendataphilly](https://www.opendataphilly.org/dataset/farmers-markets-locations/resource/d23195c1-b7ce-4ae0-a81c-51753d096c6f)

```{r warning=FALSE}
library(tidyverse)
```

```{r echo=TRUE}
markets_tidy <- readr::read_csv("../exercises/Philadelphia_Farmers_Markets.csv", )
dim(markets_tidy)

markets_base <- read.csv("../exercises/Philadelphia_Farmers_Markets.csv")
dim(markets_base)
```

## What is different?

Let's use the glimpse function from dplyr
```{r}
glimpse(markets_base)
glimpse(markets_tidy)
```

## What is different?

+ Defaults for read.csv left " ", read_csv converted to "NA"
+ Defaults for read.csv made variables factors, read_csv made character

## Fix any weirdness

+ Column names: read.csv will convert spaces to ".", read_csv will not
+ Coding of missing values: "na.strings" argument to read.csv
+ Data type doesn't match: "stringsAsFactors" argument, type.convert() function
