---
title: "The Data Cleaning Gauntlet"
author: "Alice M Walsh"
date: "8/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

Below I have assembled some data cleaning challenges! The data to clean is all available online. 

*Some of these were originally used August 2018 for an R-Ladies Philadelphia workshop*

In RStudio, you can click a green arrow in the top right corner to run the code chunk. 

# Challenge one - Ingest rectangular data

Philadelphia Farmer's Markets 
dataset sourced from https://www.opendataphilly.org/dataset/farmers-markets-locations/

This first code chunk loads the data using base R `read.csv()`. You should see new variables appear in your Global Environment. 

```{r get_data}
data_link <- 'http://data.phl.opendata.arcgis.com/datasets/0707c1f31e2446e881d680b0a5ee54bc_0.csv'
markets_base <- read.csv(data_link)
```


```{r load_packages}
# Do you want to load some packages here? Maybe tidyverse? 
# Here are the packages I loaded to complete the challenges! You made need more or different ones.

library(stringr)
library(tidyverse)
```

## Compare to readr and data.table imports

```{r}
# Add code here to load the same data (data_link) with readr::read_csv
# markets_tidy <- read_csv() ## UNCOMMENT and FINISH THE CODE
```

```{r}
# BONUS: try data.table::fread()
# markets_dt <- data.table::fread() ## UNCOMMENT and FINISH THE CODE
```


# Challenge two - Inspect your data

The first step is to inspect the data. Use 2 or three commands to look at things like...   

* What is the dimension of the data? 
* What are the variable/column names?
* What data type are the variable/columns?
* How many NA's are there per variable?

```{r inspect}
# Put your code here!

# Have you tried?
# skimr::skim()
# dplyr::glimpse()
# str()
# dim()
# colnames()
# DataExplorer::plot_intro() #Note that DataExplorer requires R >= 3.5
```

## Adjust defaults of read.csv()

There is some missing data - blank values that we would like to be represented as NA in our dataset.

Also some data has been converted to factors when it should be character.

Let's re-import the data with read.csv() with different options (to make nicer). We want the variables like NEIGHBORHOOD andd ADDRESS to be class `character`. The variable ADDRESS_NOTES should have 39 NA's.

```{r reload_data}
# Fill in the options - add stringsAsFactors, na.strings
?read.csv

# Alternatively, use your favorite package to read in the data instead of read.csv()
markets_base <- read.csv(data_link) # add here
```

# Challenge three
## Clean a character column
Let's clean up the NEIGHBORHOOD! Use all your favorite commands to look at the values in NEIGHBORHOOD and clean them up. There are some entries with an extra space. There is some inconsistent capitalization.


For example, "Center City", "Center city", and "Center City " are all different.


```{r clean_neighborhood}
# Your code here

```

## Clean by creating new variables

There are 48 values in the MONTHS variable. It looks like this variable contains information about the dates the farmer's market is open (e.g. "May - October").

If you want to plot the number of farmer's markets that are open by month, how will you do it?

Tip: Clean up the MONTHS variable and create some new variables. I also decided to create a new data frame that summarized the number of open markets by month.

```{r clean_months}
# Add code to process data here

```

```{r plot_by_month}
# Code to plot markets by month - 
# Uncomment and Modify the below ggplot2 code or start fresh with your own code

# ggplot(by_month, aes(x = month, y = number)) +
#   geom_col()
```

# Challenge four - ingest JSON and make a data.frame

I have previously downloaded data from wawa.com on locations of wawa stores.

To see what JSON looks like, paste: 'https://www.wawa.com/Handlers/LocationByLatLong.ashx?limit=50&lat=39.9526&long=-75.1652' into your web browser.

```{r}
library(jsonlite)

wawa_json <- jsonlite::fromJSON('exercises/wawa.json')
class(wawa_json)
```

wawa_json$locations has information on 50 wawa locations. However, you see that this is a nested data frame! How can we get this nice and flat!

```{r}
# Put your code here
# Try:
# jsonlite::flatten
# tidyr::unnest
# create new data.frames and then rbind, cbind?

```

## Extra challenges

That wawa data is still probably not in a good shape to analyze.
How would we make it tidy? Is it better to keep nested?

# Bonus data reshape challenge

The farmers market data is not a great example to show converting between wide and long data. If you want to practice this, here is an interesting dataset.

* Zillow makes home prices and other data available on their website: https://www.zillow.com/research/data/


```{r load_widedata}
zillow_source <- "http://files.zillowstatic.com/research/public/Zip/Zip_MedianListingPrice_AllHomes.csv"
zillow_wide <- read.csv(zillow_source)
dim(zillow_wide)
```

The data is 10165 rows and 108 columns. Let's filter to just keep Philadelphia data:

```{r filter_zillow}
# This code uses filter from dplyr package - there are other ways you could do this
zillow_wide <- filter(zillow_wide, City == "Philadelphia")
```

This data is wide - there are columns for every month with home prices (e.g. "X2018.01" and "2016.08"). Instead, we want to have one column with the prices and another with the month and year. How can we convert from wide to long?

This is roughly what I want `zillow_long` to look like:
```{r dummy_df}
fake_zillow_long <- data_frame(RegionName = c(19143, 19143, 19143, 19143, 19143),
                               City = c("Philadelphia","Philadelphia","Philadelphia","Philadelphia","Philadelphia"),
                               year = c(2010,2010,2010,2010,2010),
                               month = c(1,2,3,4,5),
                               price = c(79450, 77000, 83500, 79450, 79900))
  
fake_zillow_long
```


```{r reshape_zillow}
# Create a new dataframe, zillow_long, that contains reshaped data

```

```{r plot_zillow}
# Now we can plot the prices over time for each Philadelphia RegionName (zip code)
# Here is some example code you can modify to fit your dataset or change to make prettier

# ggplot(zillow_long, aes(x = month, y = price, group = RegionName)) + 
#   geom_line() + 
#   facet_wrap(~year)
```

