---
title: "The Data Cleaning Gauntlet"
author: "Alice M Walsh, R-Ladies Philly"
date: "8/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

Below I have assembled some data cleaning challenges! The data to clean is all available online. 

*Some of these were originally used August 2018 for an R-Ladies Philadelphia workshop*

In RStudio, you can click a green arrow in the top right corner to run the code chunk. 

# Challenge one - Ingest rectangular data

Philadelphia Farmer's Markets 
dataset sourced from https://www.opendataphilly.org/dataset/farmers-markets-locations/

This first code chunk loads the data using base R `read.csv()`. You should see new variables appear in your Global Environment. 

```{r get_data}
data_link <- 'http://data.phl.opendata.arcgis.com/datasets/0707c1f31e2446e881d680b0a5ee54bc_0.csv'
markets_base <- read.csv(data_link)
```


```{r load_packages}
# Do you want to load some packages here? Maybe tidyverse? 
# Here are the packages I loaded to complete the challenges! You made need more or different ones.

library(stringr)
library(tidyverse)
```

## Compare to readr and data.table imports

```{r}
# Add code here to load the same data (data_link) with readr::read_csv

# markets_tidy <- read_csv() ## UNCOMMENT and FINISH THE CODE
```

```{r}
# BONUS: try data.table::fread()

# markets_dt <- data.table::fread() ## UNCOMMENT and FINISH THE CODE
```


# Challenge two - Inspect your data

The first step is to inspect the data. Use two or three commands to look at things like...   

* What is the dimension of the data? 
* What are the variable/column names?
* What data type are the variable/columns?
* How many NA's are there per variable?
* Is the data "tidy"?
* What is different about data imported by read.csv and read_csv??

```{r inspect}
# Put your code here!



# Have you tried?
# skimr::skim()
# dplyr::glimpse()
# str()
# dim()
# colnames()
# DataExplorer::plot_intro() #Note that DataExplorer requires R >= 3.5
```


## Adjust defaults of read.csv(), read_csv()

There is some missing data - blank values that we would like to be represented as NA in our dataset.

Also some data has been converted to factors when it should be character.

Let's re-import the data with read.csv() with different options (to make nicer). We want the variables like NEIGHBORHOOD andd ADDRESS to be class `character`. The variable ADDRESS_NOTES should have several NA's.

```{r reload_data}
# Fill in the options - add stringsAsFactors, na.strings
?read.csv

# Alternatively, use your favorite package to read in the data 
# instead of read.csv()
markets_base <- read.csv(data_link) # add here
```

#### --- PAUSE HERE ---- ####

# Challenge three
## Clean a character column
Let's clean up the NEIGHBORHOOD! Use all your favorite commands to look at the values in NEIGHBORHOOD and clean them up. 

There might be some entries with an extra space. There is some inconsistent capitalization.

For example, "Center City", "Center city", and "Center City " are all different.


```{r clean_neighborhood}
# Create a new data.frame called markets_clean that copies the dirty data
# markets_clean <-  # ADD HERE

# Your code here
# markets_clean$NEIGHBORHOOD <- # add here
```

Bonus: What to do with missing values (NAs)? tidyr has a useful function, `replace_na()`

```{r}
?replace_na
# Try out replace_na to replace NAs in ACCEPT_SNAP_ACCESS, ACCEPT_FMNP, ACCEPT_PHILLY_FOOD_BUCKS_ with "N"

```

## Clean by creating new variables

There are 48 values in the MONTHS variable. It looks like this variable contains information about the dates the farmer's market is open (e.g. "May - October", "Open year round").

```{r}
# Currently looks like:
markets_clean[1:3, c("OBJECTID","MONTHS")]
```

If you wanted to plot the number of farmer's markets that are open by month, how would you do it? Currently, there is too much information in this single MONTHS variable.

Tip: Clean up the MONTHS variable and create some new variables. 

You want something like this:
```
  OBJECTID start_month end_month
1        1         May September
2        2         May   October
3        3     January  December
```
In this example, you have just the months in a standarized format, no dates or other values.

This value, "Wednesdays: June - August, Saturdays: May – October", is particularly painful!

```{r clean_months}
# Add code to process data here
# There are probably hundreds of approaches that could be used
# Try: separate(), mutate(), lubridate package, stringr::str_trim(), 

```


Bonus: I also decided to create a new data frame that summarized the number of open markets by month.

It will look like this:
``` r
by_month[1:5,]
     month number
1  January      6
2 February      6
3    March      6
4    April      7
5      May     26
```


```{r create_by_month}
# Code to make a new dataset that summarizes the number of open markets by month

# by_month <- #ADD HERE

```



#### --- PAUSE HERE ---- ####

# Challenge four - ingest JSON and make a data.frame

I have previously downloaded data from wawa.com on locations of wawa stores.

To see what JSON looks like, paste: 'https://www.wawa.com/Handlers/LocationByLatLong.ashx?limit=50&lat=39.9526&long=-75.1652' into your web browser.

```{r}
library(jsonlite)

wawa_json <- jsonlite::fromJSON(here::here('exercises','wawa.json'))
class(wawa_json)
```

```{r}
dplyr::glimpse(wawa_json$locations)
```


wawa_json$locations has information on 50 wawa locations. However, you see that this is a nested data frame! amenities, fuelTypes, addresses are not simple list columns.

How can we get this nice and flat!

```{r flatten_json}
# Put your code here
# Try:
# jsonlite::flatten
# tidyr::unnest
# create new data.frames and then rbind, cbind?

```


## Extra challenges

That wawa data is still probably not in a good shape to analyze.
How would we make it tidy? Is it better to keep it nested?

#### --- PAUSE HERE ---- ####

# Challenge five - pivot the "old way"

The farmers market and wawa datasets are not great examples to show converting between wide and long data. If you want to practice this, here is an interesting dataset.

* Zillow makes home prices and other data available on their website: https://www.zillow.com/research/data/


```{r load_widedata}
zillow_source <- "http://files.zillowstatic.com/research/public/Zip/Zip_MedianListingPrice_AllHomes.csv"
zillow_wide <- read_csv(zillow_source)
dim(zillow_wide)
```

The data is 10,000+ rows and 120+ columns. Let's filter to just keep Philadelphia data:

```{r filter_zillow}
# This code uses filter from dplyr package - there are other ways you could do this
zillow_wide <- filter(zillow_wide, City == "Philadelphia")
head(zillow_wide[,c(1,2,7:10)])
```

This data is wide - there are columns for every month with home prices (e.g. "2010-01" and "2016-08"). Instead, we want to have one column with the prices and another with the month and year. How can we convert from wide to long?

This is roughly what I want `zillow_long` to look like:
```{r dummy_df}
fake_zillow_long <- tibble(RegionName = c(19143, 19143, 19143, 19143, 19143),
                               City = c("Philadelphia","Philadelphia","Philadelphia","Philadelphia","Philadelphia"),
                               year = c(2010,2010,2010,2010,2010),
                               month = c(1,2,3,4,5),
                               price = c(79450, 77000, 83500, 79450, 79900))
  
fake_zillow_long
```


```{r reshape_zillow_old}
# Create a new data frame, zillow_long, that contains reshaped data
# Try using the tidyr function, gather()
?gather

# Modify below
# zillow_long <- gather(zillow_wide, 
#                       key = "year-month",
#                       value = "price",
#                       ?) # add selection of columns here!
```

```{r plot_zillow}
# Now we can plot the prices over time for each Philadelphia RegionName (zip code)
# Here is some example code you can modify to fit your dataset or change to make prettier

# ggplot(zillow_long, aes(x = month, y = price, group = RegionName)) + 
#   geom_line() + 
#   facet_wrap(~year)
```

# Challenge six - pivot the "new way"

```{r reshape_zillow_new}
# Create a new data frame, zillow_long, that contains reshaped data
?pivot_longer

# Modify below
# zillow_long <- pivot_longer(zillow_wide, 
#                       cols = , # add here!
#                       names_to = "year-month",
#                       values_to = "price")
```

OK, so now the pivot_wider() function is the opposite!

Let's practice by pivoting zillow_long to a new wide table where each month has it's own column. It will look like this (with some additional columns and rows):

```{r dummy_df_2}
fake_zillow_wide2 <- tibble(RegionName = c(19143, 19143, 19143),
                            City = c("Philadelphia","Philadelphia","Philadelphia"),
                            year = c(2010,2011,2012),
                            price_01 = c(180000, 171400, 164900),
                            price_02 = c(179999, 169450, 165000),
                            price_03 = c(179900, 174900, 165000))
  
fake_zillow_wide2
```

```{r back_to_wide}
# Create a new data frame, zillow_long, that contains reshaped data
?pivot_wider

# Modify below
# zillow_wide_2 <- pivot_wider(zillow_long,
#                       names_from = ,
#                       values_from = ,
#                       names_prefix = )
```

#### --- PAUSE HERE ---- ####
# Reproducible examples with reprex

https://github.com/tidyverse/reprex

Here are some examples to create a reprex
```{r}
library(reprex)
```

```{r}
# default is for a github issue
# copy (ctrl+c) the below lines to the clipboard
bad_data <- data.frame(money = c(10,20,30,30),
                       name = c("alice","alice","dave","dave"),
                       weekday = c("monday","tuesday","monday","tuesday"))
bad_data
good_data <- tidyr::spread(bad_data, key=name, value=money)
good_data
# then run this
reprex()
# can now paste on github or stackoverflow
```

You don't want to post a question using data you have loaded, but no one else has.
Hence, creating the dummy `bad_data` data.frame.
Another method is to just use some slice of your real data if it is shareable.

```{r}
# try dput()
dput(df[1:6,1:6])
```

