---
title: "The Data Cleaning Gauntlet"
subtitle: "Possible Answers"
author: "Alice M Walsh, R-Ladies Philly"
date: "3/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Solutions

# Challenge one - Ingest rectangular data

Philadelphia Farmer's Markets 
dataset sourced from https://www.opendataphilly.org/dataset/farmers-markets-locations/

This first code chunk loads the data using base R `read.csv()`. You should see new variables appear in your Global Environment. 

```{r get_data}
data_link <- 'http://data.phl.opendata.arcgis.com/datasets/0707c1f31e2446e881d680b0a5ee54bc_0.csv'
markets_base <- read.csv(data_link)
```


```{r load_packages}
# Do you want to load some packages here? Maybe tidyverse? 
# Here are the packages I loaded to complete the challenges! You made need more or different ones.

library(stringr)
library(lubridate)
library(tidyverse)
```

## Compare to readr and data.table imports

```{r}
# Add code here to load the same data (data_link) with readr::read_csv

markets_tidy <- read_csv(data_link) ## UNCOMMENT and FINISH THE CODE
```

```{r}
# BONUS: try data.table::fread()

markets_dt <- data.table::fread(data_link) ## UNCOMMENT and FINISH THE CODE
```

# Challenge two - Inspect your data

```{r inspect}
# Here are some useful data inspection functions!
dim(markets_tidy)
names(markets_tidy)
str(markets_tidy)
summary(markets_tidy)
colSums(is.na(markets_tidy))
# Using dplyr
glimpse(markets_tidy)
# Use lapply to check class or number of NA?
lapply(markets_tidy, class)
lapply(markets_tidy, function(x) sum(is.na(x)))

## What is different?

# + Defaults for read.csv left " ", read_csv converted to "NA"
# + Defaults for read.csv made variables factors, read_csv made character
# + Defaults for read.csv left white space, read_csv trimmed it

## Fix any weirdness

# + Column names: read.csv will convert spaces to ".", read_csv will not
# + Coding of missing values: "na.strings" argument to read.csv
# + Data type doesn't match: "stringsAsFactors" argument, type.convert() function

```

## Adjust defaults of read.csv(), read_csv()

```{r reload_data}
# Fill in the options - add stringsAsFactors, na.strings
# ?read.csv

markets_base <- read.csv(data_link, 
                         stringsAsFactors = FALSE, 
                         strip.white = TRUE,
                         na.strings=c("NA","NaN", " ", ""))
# Alternatively, use your favorite package to read in the data instead of read.csv()
markets <- read_csv(data_link, na = c("NA","NaN", " ", ""))
```

# Challenge three
## Clean a character column

```{r clean_neighborhood}
# Create a new data.frame called markets_clean that copies the dirty data
markets_clean <- markets

# One way
markets_clean <- markets_clean %>% #note this saves the changes to the dataframe
  # stringr package str_trim() - dont need if done on import
  mutate(NEIGHBORHOOD = stringr::str_trim(NEIGHBORHOOD)) %>% 
  # str_to_title - makes title case
  mutate(NEIGHBORHOOD = stringr::str_to_title(NEIGHBORHOOD))

# Check
unique(markets_clean$NEIGHBORHOOD)
# or 
markets_clean %>% count(NEIGHBORHOOD)

```

Bonus: What to do with missing values (NAs)? tidyr has a useful function, `replace_na()`

```{r}
# ?replace_na
# Try out replace_na to replace NAs in ACCEPT_SNAP_ACCESS, ACCEPT_FMNP, ACCEPT_PHILLY_FOOD_BUCKS_ with "N"

markets_clean <- markets_clean %>%   
  replace_na(list(ACCEPT_SNAP_ACCESS = "N",
                  ACCEPT_FMNP = "N",
                  ACCEPT_PHILLY_FOOD_BUCKS_ = "N"))
```

## Clean by creating new variables

```{r clean_months_one}
# This is one way... 

# Create a factor with the FULL months in order (will be useful later)
# base R has `month.name`,`month.abb`
all_months <- factor(month.name, levels= month.name)
# do same for abvreviations
all_months_abb <- factor(month.abb, levels= month.abb)

# This cleanup with pipes (%>%) does several operations
markets_clean <- markets_clean %>% 
  # Manual clean-up
  mutate(MONTHS = sub("TBD|Late |end of |Mid-|mid-|middle of |,","", MONTHS)) %>% 
  mutate(MONTHS = sub("Thanksgiving","November", MONTHS)) %>% 
  # Change all to uppercase
  mutate(MONTHS = str_to_title(MONTHS)) 
# Search for any months present in MONTHS - convert to numbers and then find min and max
markets_clean$month_n_min <- sapply(markets_clean$MONTHS,
      function(m) min(as.numeric(all_months[str_detect(m, as.character(all_months))]),
                      as.numeric(all_months_abb[str_detect(m, as.character(all_months_abb))])))

markets_clean$month_n_max <- sapply(markets_clean$MONTHS,
      function(m) max(as.numeric(all_months[str_detect(m, as.character(all_months))]),
                    as.numeric(all_months_abb[str_detect(m, as.character(all_months_abb))])))

markets_clean <- markets_clean %>% 
  # Special case of year round - make start 1 (Jan), end 12 (Dec)
  mutate(month_n_min = if_else(grepl("Round",MONTHS), 1, month_n_min),
         month_n_max = if_else(grepl("Round",MONTHS), 12, month_n_max)) %>% 
  # Replace Inf, -Inf with NA
  mutate_at(vars(starts_with("month_n")), list(~na_if(abs(.), Inf)))

# The last line above is pretty hard to understand, 
# here is another way (non-tidyverse) to do it
# markets$month_n_min[is.infinite(markets$month_n_min)] <- NA
# markets$month_n_max[is.infinite(markets$month_n_max)] <- NA

```

```{r plot_by_month}

# Create a summarization based on the numeric start and end months

by_month <- data.frame(month = all_months,
                       number = sapply(1:12,
                                function(x) 
                                sum(markets_clean$month_n_min <= x & markets_clean$month_n_max >= x, 
                                    na.rm = TRUE)))
# Code to plot markets by month
ggplot(by_month, aes(x = month, y = number)) +
  geom_col()
```


```{r clean_months_two, eval=FALSE}
# Here is a different way to approach challenge four
markets2 <- markets
# Uses lubridate
unique(markets2$MONTHS)

# split by hyphen or dash...
markets2$START_DATE <- str_split_fixed(markets2$MONTHS, "\\p{Pd}", n = 2)[,1]
markets2$END_DATE <- str_split_fixed(markets2$MONTHS, "\\p{Pd}", n = 2)[,2]

# for "year round", start date is January and end date is December
year_round_index <- grepl("year round", markets2$MONTHS, ignore.case = TRUE)
markets2$START_DATE[year_round_index] = "2018-01-01" # year required for lubridate so using 2018, is there a better way?
markets2$END_DATE[year_round_index] = "2018-12-01"

# grep for months within text before first hyphen/dash
# if I had more time could make this a loop through abbreviated months of the year?
unique(markets2$START_DATE)
january_index <- grepl("January", markets2$START_DATE, ignore.case = TRUE)
markets2$START_DATE[january_index] = "2018-04-01"
april_index <- grepl("April", markets2$START_DATE, ignore.case = TRUE)
markets2$START_DATE[april_index] = "2018-04-01"
may_index <- grepl("May", markets2$START_DATE, ignore.case = TRUE)
markets2$START_DATE[may_index] = "2018-05-01"
june_index <- grepl("June", markets2$START_DATE, ignore.case = TRUE)
markets2$START_DATE[june_index] = "2018-06-01"
july_index <- grepl("July", markets2$START_DATE, ignore.case = TRUE)
markets2$START_DATE[july_index] = "2018-07-01"


# grep for months within text after first hyphen/dash...
unique(markets2$END_DATE)
aug_index <- grepl("Aug", markets2$END_DATE, ignore.case = TRUE)
markets2$END_DATE[aug_index] = "2018-08-01"
sept_index <- grepl("Sept", markets2$END_DATE, ignore.case = TRUE)
markets2$END_DATE[sept_index] = "2018-09-01"
oct_index <- grepl("Oct", markets2$END_DATE, ignore.case = TRUE)
markets2$END_DATE[oct_index] = "2018-10-01"
nov_index <- grepl("Nov|Thanksgiving", markets2$END_DATE, ignore.case = TRUE)
markets2$END_DATE[nov_index] = "2018-11-01"

# check start and end dates
unique(markets2$START_DATE)
unique(markets2$END_DATE)


# We have messed up some months...
dplyr::select(markets2, MONTHS, START_DATE, END_DATE) %>% slice(47)


# use lubridate to create variable for date interval the market is open
markets2$INT_OPEN <- interval(markets2$START_DATE, markets2$END_DATE)
# Remove any NA in INT_OPEN - just have year round
markets2$INT_OPEN[is.na(markets2$INT_OPEN)] <- interval("2018-01-01", "2018-12-01")
# I was trying to find a way of plotting the interval directly, but didn't find anything

# build data frame with row for each month, first col = month number, second col = num markets open, third col = iso date for first of month (2018)
# sum() is counting the number of rows for which the specified date is within the date interval

markets_open_count <- as.data.frame(
  rbind(
    "01" = c(01, sum(
      ymd("2018-01-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-01-01"),
    "02" =  c(02, sum(
      ymd("2018-02-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-02-01"),
    "03" =  c(03, sum(
      ymd("2018-03-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-03-01"),
    "04" =  c(04, sum(
      ymd("2018-04-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-04-01"),
    "05" =  c(05, sum(
      ymd("2018-05-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-05-01"),
    "06" =  c(06, sum(
      ymd("2018-06-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-06-01"),
    "07" =  c(07, sum(
      ymd("2018-07-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-07-01"),
    "08" =  c(08, sum(
      ymd("2018-08-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-08-01"),
    "09" =  c(09, sum(
      ymd("2018-09-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-09-01"),
    "10" =  c(10, sum(
      ymd("2018-10-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-10-01"),
    "11" =  c(11, sum(
      ymd("2018-11-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-11-01"),
    "12" =  c(12, sum(
      ymd("2018-12-01") %within% markets2$INT_OPEN == TRUE
    ), "2018-12-01")
  )
)

colnames(markets_open_count) <- c("month", "count", "date")
markets_open_count$month <- as.factor(markets_open_count$month)
markets_open_count$count <- as.numeric(as.character(markets_open_count$count))
markets_open_count
```


```{r plot_by_month_two, eval=FALSE}
# Code to plot markets by month - 

ggplot(markets_open_count, aes(x = month(date, label = TRUE), y = count)) +
  geom_col() +
  scale_x_discrete(limits = month.abb)
```

# Challenge four

```{r load_json}
library(jsonlite)

wawa_json <- jsonlite::fromJSON(here::here('exercises','wawa.json'))
class(wawa_json)
```

```{r}
dplyr::glimpse(wawa_json$locations)
```

```{r flatten_json}
# One way
flat_wawa <- wawa_json$locations

# amenitites can get added with simple jsonlite::flatten()
# dim(flat_wawa$amenities)
flat_wawa <- flatten(flat_wawa)

# addresses has two rows x 6 cols per row, can use tidyr::unnest()
# dim(flat_wawa$addresses[[1]])
flat_wawa <- unnest(flat_wawa, cols = addresses)

# fuelTypes has data frame for locations with gas
#   description price currency
# 1    Unleaded 2.419      USD
# 2        Plus 2.739      USD
# 3     Premium 2.879      USD
# 4      Diesel 2.999      USD
# 5       Super 2.919      USD

flat_wawa <- unnest(flat_wawa, cols = fuelTypes, keep_empty = TRUE)

dim(flat_wawa)
```


# Challenge five - pivot the "old way"

```{r load_widedata}
zillow_source <- "http://files.zillowstatic.com/research/public/Zip/Zip_MedianListingPrice_AllHomes.csv"
zillow_wide <- read_csv(zillow_source)
dim(zillow_wide)
```

The data is 10,000+ rows and 120+ columns. Let's filter to just keep Philadelphia data:

```{r filter_zillow}
# This code uses filter from dplyr package - there are other ways you could do this
zillow_wide <- filter(zillow_wide, City == "Philadelphia")
head(zillow_wide[,c(1,2,7:10)])
```

This data is wide - there are columns for every month with home prices (e.g. "2010-01" and "2016-08"). Instead, we want to have one column with the prices and another with the month and year. How can we convert from wide to long?

This is roughly what I want `zillow_long` to look like:
```{r dummy_df}
fake_zillow_long <- tibble(RegionName = c(19143, 19143, 19143, 19143, 19143),
                               City = c("Philadelphia","Philadelphia","Philadelphia","Philadelphia","Philadelphia"),
                               year = c(2010,2010,2010,2010,2010),
                               month = c(1,2,3,4,5),
                               price = c(79450, 77000, 83500, 79450, 79900))
  
fake_zillow_long
```


```{r reshape_zillow_old}
# Create a new data frame, zillow_long, that contains reshaped data
# Try using the tidyr function, gather()
# ?gather

# One solution (nice b/c dont need to know all col names)
zillow_long <- gather(zillow_wide, 
                      key = "year-month", #bad col name!
                      value = "price",
                      starts_with("20")) %>% 
  separate(`year-month`, 
           into = c("year","month"))

# Basically the same but specify column range instead of starts_with
# also dont autofill NA
zillow_long <- gather(zillow_wide, 
                      key = "year-month", #bad col name!
                      value = "price",
                      `2010-01`:`2019-12`,
                      na.rm = TRUE) %>% 
  separate(`year-month`, 
           into = c("year","month"), 
           convert = TRUE) #if you want new year and month cols to be numeric
```

```{r plot_zillow}
# Now we can plot the prices over time for each Philadelphia RegionName (zip code)
# Here is some example code you can modify to fit your dataset or change to make prettier

ggplot(zillow_long, aes(x = month, y = price, group = RegionName)) +
  geom_line() +
  facet_wrap(~year)
```

# Challenge six - pivot the "new way"

```{r reshape_zillow_new}
# Create a new data frame, zillow_long, that contains reshaped data
# ?pivot_longer

# One way
zillow_long <- pivot_longer(zillow_wide, 
                      cols = starts_with("20"),
                      names_to = "year-month",
                      values_to = "price") %>% 
  separate(`year-month`, 
           into = c("year","month"))

# With some other options
zillow_long <- pivot_longer(zillow_wide, 
                      cols = starts_with("20"),
                      names_to = "year-month",
                      values_to = "price",
                      values_drop_na = TRUE) %>% 
  separate(`year-month`, 
           into = c("year","month"))
```

```{r back_to_wide}
# Create a new data frame, zillow_long, that contains reshaped data
# ?pivot_wider

# Modify below
zillow_wide_2 <- pivot_wider(zillow_long,
                      names_from = "month",
                      values_from = "price",
                      names_prefix = "price_")
```


